The imprint of global magma oceans on exoplanet demographics

### 2023-04-09 src/tex/ms.tex
6.6.3

\subsubsection{Simplifications in the atmosphere model}
\todo[inline]{TIM: your expertise will be valuable here.\\
Mention we only consider instellation-induced rgh.\\
- we ignore water loss by H$_2$O photolysis in the upper atmosphere and subsequent H escape, which would eventually cool even planets within the runaway greenhouse regime~\citep{Lichtenberg2022}.\\
- we ignore tidal heating, which could extend the magma ocean phase of close-in planets (and change their orbits via tidal orbital evolution)( Jackson et al., 2008; Barnes et al., 2013 (?)).\\
- assumption of a single albedo. ``if there is a range of albedos, will that wash away the signal?''
}
As illustrated in Fig.~\ref{fig:radiusevolution}, evolutionary processes and a background population of planets unable to sustain a steam atmosphere complicate the classification of a planet into the runaway or sub-runaway category.
For instance, both a Hydrogen/Helium-dominated ``sub-Neptune'' and a terrestrial planet currently hosting a steam atmosphere will appear as a planet in our runaway greenhouse category, diluting the expected demographic signature.
Further, a planet in a runaway greenhouse state will eventually lose its steam atmosphere and move to the sub-runaway greenhouse category on a timescale of $\SI{e8}{\year}$ for a planet with one Earth ocean of water~\citep[][]{Watson1981}.
We marginalized over these degeneracies by introducing the dilution factor $f_\mathrm{rgh}$, but in a real survey contextual information about a planet's age or likely formation history could be taken into account.

>>>>

\subsubsection{Simplifications in the atmosphere model}
%\todo[inline]{TIM: your expertise will be valuable here.\\
%Mention we only consider instellation-induced rgh.\\
%- we ignore water loss by H$_2$O photolysis in the upper atmosphere and subsequent H escape, which would eventually cool even planets %within the runaway greenhouse regime~\citep{Lichtenberg2022}.\\
%- we ignore tidal heating, which could extend the magma ocean phase of close-in planets (and change their orbits via tidal orbital evolution)( Jackson et al., 2008; Barnes et al., 2013 (?)).\\
%- assumption of a single albedo. ``if there is a range of albedos, will that wash away the signal?''
%}
% >>>>>>>>>>
% TL: I just wrote a review article about this exact problem (magma ocean atmosphere exchange) with Yamila Miguel, hopefully soon on arXiv. Honestly, I think we keep it short here, otherwise the paper will get way too long and diluted. I referenced the latest reviews on this now, and we can cite Lichtenberg & Miguel 23 as soon as it's on arXiv for a much much longer discussion.
% >>>>>>>>>>
As illustrated in Fig.~\ref{fig:radiusevolution}, evolutionary processes and a background population of planets unable to sustain a steam atmosphere complicate the classification of a planet into the super- or sub-runaway category. For instance, both a Hydrogen/Helium-dominated ``sub-Neptune'' and a terrestrial planet currently hosting a steam atmosphere will appear as a planet in our runaway greenhouse category, diluting the expected demographic signature. Further, a planet in a runaway greenhouse state will eventually lose its steam atmosphere and move to the sub-runaway greenhouse category on a timescale of $\SI{e8}{\year}$ for a planet with one Earth ocean of water~\citep[][]{Watson1981}. Developing more sophisticated models of the coupled interaction between atmosphere and interior for an arbitrary planetary composition remains a long-term goal of the community \citep{2021SSRv..217...22G,Lichtenberg2022,2022ARA&A..60..159W}.
We marginalized over these degeneracies by introducing the dilution factor $f_\mathrm{rgh}$, but in a real survey contextual information about a planet's age or likely formation history could be taken into account.
